---
title: "Athletics Analysis"
output:
  html_document:
    df_print: paged
---


```{r set-up}
library(tidyverse)
library(lubridate)

athletes <- read_csv("https://raw.githubusercontent.com/rdpapworth/Athletics/master/data/athletes.txt",
                         col_types = cols(dob = col_date(format = "%Y%m%d")))

performances <- read_csv("https://raw.githubusercontent.com/rdpapworth/Athletics/master/data/performances.txt",
                         col_types = cols(dop = col_date(format = "%Y%m%d")))

results <- inner_join(performances, athletes, by=c("athlete_id"="id"))
```

## Is there a Relative Age Effect?

The results are for top ranked athletes nationally. Since age-groups are based on a September cut-off, there seems to be a selection bias.

```{r month_of_birth}
athletes %>%
  mutate(mob = month(dob, label = T)) %>%
  ggplot(aes(x = mob)) +
  geom_bar()
```

## Does it differ by event?

Quick visualisation suggests effect is apparent in all the available events (I thought I'd downloaded more. Will check archived version of download code)

Is there a way to identify whether one event is more impacted than another?
```{r by_event}
results %>%
  mutate(mob = month(dob, label = T)) %>%
  ggplot(aes(x = mob)) +
  geom_bar() +
  facet_grid(event_name ~ .)
```


## Have performances improved?
```{r yoy_improvement}

ts <- results %>%
  filter(event_name %in% c('400') & as.numeric(dop - dob)/365.25 >= 18.0) %>%
  mutate(yr = lubridate::year(dop), performance = as.numeric(performance)) %>%
  group_by(sex, yr) %>%
  top_n(-20,performance) %>%
  summarise(ave_performance = mean(performance)) %>%
  arrange(yr)

ggplot(ts, aes(x=yr)) + 
  geom_line(aes(y=ave_performance, col=sex)) 
```
Seems to be a general downward trend at the top. Will this hold at middle/botom. Need to figure out how to properly plot ts (breaks and labels). Would be good to show range of results (boxplot with year on x and perofrmance on y?) 